Measuring Political Freedom Around the World

### 2020-11-15 slides.Rmd
added theme

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(patchwork)

theme_set(theme_light(base_size = 30))
options(htmltools.dir.version = FALSE, echo = FALSE)
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 8)

@@

@@

```{r}
set.seed(69)
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")
t2[3,2] <- median(t1[,2], na.rm = TRUE)
colnames(t2) <- c("Obs. Number", "Data Point")
knitr::kable(list(t1, t2), caption = "Illustrated Example")
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(patchwork)
library(xaringanthemer)

style_mono_accent_inverse(base_color = "#3C989E")
theme_set(theme_light(base_size = 30))
options(htmltools.dir.version = FALSE, echo = FALSE)
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 8)

@@

```{r xaringan-themer, include=FALSE, warning=FALSE}

@@

```{r}
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")
t2[3,2] <- median(t1[,2], na.rm = TRUE)
colnames(t2) <- c("Obs. Number", "Data Point")
knitr::kable(list(t1, t2), caption = "Illustrated Example")
```

### 2020-11-15 themed_slides.Rmd
added theme

---
title: "Measuring Political Freedom Around the World"
subtitle: "Sabrina Ball, Joe Despres, Joe Sigler, Jacob Haywood"
date: "Last compiled `r format(Sys.time(), '%H:%M')`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

```{r xaringan-themer, include=FALSE, warning=FALSE}

@@

---
class: center
# OECD

@@

```{r}
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")
t2[3,2] <- median(t1[,2], na.rm = TRUE)
colnames(t2) <- c("Obs. Number", "Data Point")
knitr::kable(list(t1, t2), caption = "Illustrated Example")
```

@@

data %>% 
  group_by(polrights_fh) %>% 
  add_count(polrights_fh) %>% 
  ggplot(aes(x = polrights_fh, fill = factor(polrights_fh)), color = "black") + 
  geom_bar() + 
  geom_text(data = lab, aes(x = polrights_fh, y = n, label = n), vjust = -.75) + 
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_x_continuous(breaks = 1:7) + 
  theme_test(base_size = 30) + 
  theme(legend.position = "none") + 
  labs(y = "Number", x = "Political Freedom", title = "Distribution of Political Freedom Within the OECD")
```

@@

subsets %>% 
  ggplot(aes(x = df, y = adj.r.squared)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, adj), se = FALSE, span = 0.5, color = "Green") + 
  labs(title = "", x = "Parameters", y = "Adjusted R Squared") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30)
```

@@

subsets %>% 
  ggplot(aes(x = df, y = AIC)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, AIC), se = FALSE, span = 0.5, color = "red") + 
  labs(title = "", x = "Parameters", y = "AIC") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30)
```

@@

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"))  %>% 
  row_spec(2, color = "white", background = "#D7261E")
```

@@

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(3, color = "white", background = "#D7261E")
```

@@

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(4, color = "white", background = "#D7261E")
```

@@

class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(5, color = "white", background = "#D7261E")
```

@@

class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(6, color = "white", background = "#D7261E")
```

@@

hist <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(.resid)) + 
  geom_histogram(bins = 100, fill = "black") +
  scale_x_continuous(limits = c(-4,4)) + labs(y = "Density", x = "Residual", title = "Histogram")

qq <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(sample = .resid)) + 
  stat_qq() +
  stat_qq_line() + 
scale_x_continuous(limits = c(-4,4)) + 
scale_y_continuous(limits = c(-4,4)) + labs(title = "QQ Plot of Residuals ", x = "Residual", y = "Expected")
```

@@

```{r}
fit_df5 %>% 
  broom::augment() %>% 
  mutate(yhat = predict(fit_df5),
         resid = resid(fit_df5)) %>% 
  ggplot() + 
  geom_point(aes(x = yhat, y = resid)) + 
  geom_hline(yintercept = 0, color = "red") + 
  scale_y_continuous(limits = c(-4,4)) + 
  labs(y = "Residual", x = "Predicted Value")
```

@@

```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Included Outliers") %>% 
  row_spec(3, color = "white", background = "#D7261E")
```

@@

```{r}
data %>% 
  filter(polrights_fh >= 4) %>% 
  lm(model_df5_formula, .) %>% 
  broom::tidy() %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Removed Outliers") %>% 
  row_spec(3, color = "white", background = "#D7261E")
```

---
```{r, animation.hook="gifski", interval = .1}
for(i in 1:(nrow(data))/5){
plot(data$electricity_use_per_person[1:i*5], resid(fit_df5)[1:i*5], xlim=c(0,6000), ylim = c(-4, 1.25))
}

@@

- We identified these outliers.
- confirmed they are not errors. 
- decided to keep them because They are indicating a substancial loss in political freedom which falls outside of the scope of the model. However, is important for our reserach Q.
***
> A safe rule frequently suggested is to discard an outlier only if there is direct evidence that it represents an error in recording, a miscalculation, a malfunctioning of equipment, or a similar type of circumstance. -Chapter 3 p.108

# Brown Forsythe Test

>>>>

---
title: "Measuring Political Freedom Around the World"
subtitle: "Sabrina Ball, Joe Despres, Joe Sigler, Jacob Haywood"
date: "Last compiled `r format(Sys.time(), '%H:%M')`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

@@

class: center, white

# OECD

@@

```{r}
set.seed(69)
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")
t2[3,2] <- median(t1[,2], na.rm = TRUE)
colnames(t2) <- c("Obs. Number", "Data Point")
knitr::kable(list(t1, t2), caption = "Illustrated Example")
```

@@

data %>% 
  group_by(polrights_fh) %>% 
  add_count(polrights_fh) %>% 
  ggplot(aes(x = polrights_fh, fill = factor(polrights_fh)), color = "#3C989E") + 
  geom_bar() + 
  geom_text(data = lab, aes(x = polrights_fh, y = n, label = n), color = "#3C989E", size = 8) + 
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_x_continuous(breaks = 1:7) + 
  theme_test(base_size = 30) + 
  theme_xaringan() + 
  theme(legend.position = "none") + 
  labs(y = "Number", x = "Political Freedom", title = "Distribution of Political Freedom Within the OECD")
```

@@

subsets %>% 
  ggplot(aes(x = df, y = adj.r.squared)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, adj), se = FALSE, span = 0.5, color = "red") + 
  labs(title = "", x = "Parameters", y = "Adjusted R Squared") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30) + 
  theme_xaringan() 
```

@@

subsets %>% 
  ggplot(aes(x = df, y = AIC)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, AIC), se = FALSE, span = 0.5, color = "red") + 
  labs(title = "", x = "Parameters", y = "AIC") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30) + 
  theme_xaringan() 
```

@@

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"))  %>% 
  row_spec(2, color = "white", background = "#3C989E")
```

@@

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(3, color = "white", background = "#3C989E")
```

@@

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(4, color = "white", background = "#3C989E")
```

@@

class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(5, color = "white", background = "#3C989E")
```

@@

class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(6, color = "white", background = "#3C989E")
```

@@

hist <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(.resid)) + 
  geom_histogram(bins = 100) +
  scale_x_continuous(limits = c(-4,4)) + labs(y = "Density", x = "Residual", title = "Histogram") + theme_xaringan()

qq <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(sample = .resid)) + 
  stat_qq() +
  stat_qq_line() + 
scale_x_continuous(limits = c(-4,4)) + 
scale_y_continuous(limits = c(-4,4)) + labs(title = "QQ Plot of Residuals ", x = "Residual", y = "Expected") + theme_xaringan() 
```

@@

```{r}
fit_df5 %>% 
  broom::augment() %>% 
  mutate(yhat = predict(fit_df5),
         resid = resid(fit_df5)) %>% 
  ggplot() + 
  geom_point(aes(x = yhat, y = resid)) + 
  geom_hline(yintercept = 0, color = "red") + 
  scale_y_continuous(limits = c(-4,4)) + 
  theme_xaringan() +
  labs(y = "Residual", x = "Predicted Value")
```

@@

```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Included Outliers") %>% 
  row_spec(3, color = "white", background = "#3C989E")
```

@@

```{r}
data %>% 
  filter(polrights_fh >= 4) %>% 
  lm(model_df5_formula, .) %>% 
  broom::tidy() %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Removed Outliers") %>% 
  row_spec(3, color = "white", background = "#3C989E")
```

@@

- We identified these outliers.
- confirmed they are not errors. 
- decided to keep them because They are indicating a substancial loss in political freedom which falls outside of the scope of the model. However, is important for our reserach Q.
***
> "A safe rule frequently suggested is to discard an outlier only if there is direct evidence that it represents an error in recording, a miscalculation, a malfunctioning of equipment, or a similar type of circumstance". -Chapter 3 p.108

---
class: center
# Brown Forsythe Test
